---
title: "Integration EDA"
output: html_notebook
---


```{r}
library(Seurat)
library(tidyverse)
library(SingleCellExperiment)
library(cowplot)
library(MultiAssayExperiment)
library(reshape2)
source("~/multiOmic_benchmark/selectFeatures.R")
```

Load integration output
```{r}
int.cca <- readRDS("~/models/integrateCCA_F74_SCElist_20191017.RDS")
int.liger <- readRDS("~/models/integrateliger_F74_SCElist_20191017.RDS")

int.features <- int.cca$misc$transfer.anchors@anchor.features

```

<!-- ```{r} -->
<!-- pseudocells.cca <- int.cca$intOut[["integrated.RNA"]]  -->
<!-- pseudocells.liger <- int.liger$intOut[["integrated.RNA"]]  -->
<!-- normcounts(int.liger$intOut[["integrated.RNA"]][,int.liger$intOut$tech=="RNA"])  <- normcounts(int.liger$intOut[["RNA"]][int.features,] ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # sce.list <- readRDS("my_data/integrated_thymus/F74_SCElist_20191017.RDS") -->
<!-- # small.sce.list <- map(sce.list, ~ .x[, sample(1:6000, size = 2000)]) -->
<!-- # integrate_features <- HVG_Seurat(sce.list$RNA) -->
<!-- # # transfer.anchors <- readRDS("models/F74_seuratCCA_20191017.RDS") -->
<!-- # small.sce.list.integration <- integrate_seuratCCA(small.sce.list, integrate_features = integrate_features) -->
<!-- plotUMAP <- function(pseudocells){ -->
<!--   pseudocells <- scater::runPCA(pseudocells, ncomponents=50, exprs_values="normcounts", scale=TRUE, feature_set = int.features) -->
<!--   pseudocells <- scater::runUMAP(pseudocells, use_dimred = "PCA", feature_set = integrate_features, n_neighbors = 30) -->
<!--   umap.sce <- SingleCellExperiment::reducedDim(pseudocells, "UMAP") -->
<!--   pca.sce <- SingleCellExperiment::reducedDim(pseudocells, "PCA") -->
<!--   # pca.sce %>% as_tibble(rownames="cell") %>% -->
<!--   #   left_join(as_tibble(colData(pseudocells), rownames="cell")) %>%  -->
<!--   #   ggplot(aes(PC1, PC2, color=tech)) + -->
<!--   #   geom_point(alpha=0.2) -->
<!--   umap.sce %>% as_tibble(rownames="cell") %>% -->
<!--     left_join(as_tibble(colData(pseudocells), rownames="cell")) %>%  -->
<!--     ggplot(aes(V1, V2, color=tech)) + -->
<!--     geom_point(alpha=0.2) -->

<!-- } -->

<!-- plotUMAP(pseudocells.cca) -->
<!-- plotUMAP(pseudocells.liger) -->
<!-- ``` -->
<!-- # ```{r} -->
<!-- # pseudocells.liger <- scater::runPCA(pseudocells.liger, ncomponents=50, exprs_values="normcounts", scale=TRUE, feature_set = int.features) -->
<!-- #  -->
<!-- # pca.sce <- SingleCellExperiment::reducedDim(pseudocells.liger, "PCA") -->
<!-- # pca.sce %>% as_tibble(rownames="cell") %>% -->
<!-- #   left_join(as_tibble(colData(pseudocells.liger), rownames="cell")) %>% -->
<!-- #   ggplot(aes(PC1, PC2, color=tech)) + -->
<!-- #   geom_point(alpha=0.2) -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->

```{r, fig.width=12, fig.height=5}
# rna.seu <- as.Seurat(int.cca$intOut[["RNA"]])
# atac.seu <- as.Seurat(int.cca$intOut[["ATAC"]], assay="ATAC")
# 
# imputed.assay <- CreateAssayObject(assay(int.cca$intOut[["integrated.RNA"]][,which(int.cca$intOut$tech=="ATAC")], 'normcounts'))
# atac.seu[["RNA"]] <- imputed.assay
# atac.seu@active.assay <- "RNA"
# coembed <- merge(x = rna.seu@assays$RNA, y = imputed.assay)
intOutput2seurat <- function(intOut, int.features){
  int.seu <- as.Seurat(intOut[["integrated.RNA"]], counts=NULL, data="normcounts")
  int.seu@assays$RNA@var.features <- int.features
  return(int.seu)
  }

int.cca.seu <- intOutput2seurat(int.cca$intOut, int.features)

int.cca.seu <- ScaleData(int.cca.seu, features = int.features)
int.cca.seu <- RunPCA(int.cca.seu)
int.cca.seu <- RunUMAP(int.cca.seu, reduction="pca", dims=1:50)

int.liger.seu <- intOutput2seurat(int.liger$intOut, int.features = int.features)

int.liger.seu <- ScaleData(int.liger.seu, features = int.features)
int.liger.seu <- RunPCA(int.liger.seu)
int.liger.seu <- RunUMAP(int.liger.seu, reduction="pca", dims=1:50)

p1 <- DimPlot(int.liger.seu, reduction = "umap", group.by="tech") + ggtitle("LIGER")
p2 <- DimPlot(int.cca.seu, reduction = "umap", group.by="tech") + ggtitle("CCA")

ggpubr::ggarrange(p1, p2)
```

### LIGER factor model

Testing Liger integration as the authors
```{r}
ligerH.umap <- uwot::umap(int.liger$misc$H.norm, pca = NULL, verbose = TRUE)
ligerHraw.umap <- uwot::umap(purrr::reduce(int.liger$misc$H, rbind), pca = NULL, verbose = TRUE)

pl_hnorm <- as_tibble(ligerH.umap) %>%
  mutate(cell=rownames(int.liger$misc$H.norm)) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  ggplot(aes(V1, V2, color=tech)) +
  geom_point(size=0.5, alpha=0.4)
pl_h <- as_tibble(ligerHraw.umap) %>%
  mutate(cell=rownames(int.liger$misc$H.norm)) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  ggplot(aes(V1, V2, color=tech)) +
  geom_point(size=0.5, alpha=0.4)

ggpubr::ggarrange(pl_h + ggtitle("Liger H"), pl_hnorm + ggtitle("Liger H_norm"), common.legend = TRUE)
```

Inspection of NMF factor matrix before and after normalization
```{r}
smp.cells.rna <- sample(colnames(int.liger$intOut[["RNA"]]), 500)
smp.cells.atac <- sample(colnames(int.liger$intOut[["ATAC"]]), 500)

H <- int.liger[["misc"]]$H
H.norm <- int.liger[["misc"]]$H.norm

rbind(H$RNA[smp.cells.rna, ], H$ATAC[smp.cells.atac, ]) %>% 
  melt(varnames=c("cell", "factor")) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  group_by(factor) %>%
  mutate(rank=rank(value)) %>%
  ungroup() %>%
  dplyr::filter(factor< 12) %>%
  ggplot(aes(tech, value)) +
  ggbeeswarm::geom_quasirandom(aes(color=tech), alpha=0.4)  +
  facet_wrap(factor~., scales="free_y") +
  ggtitle("H")
  
H.norm[c(smp.cells.atac, smp.cells.rna),] %>% 
  melt(varnames=c("cell", "factor")) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  group_by(factor) %>%
  mutate(rank=rank(value)) %>%
  ungroup() %>%
  dplyr::filter(factor< 12) %>%
  ggplot(aes(tech, value)) +
  ggbeeswarm::geom_quasirandom(aes(color=tech), alpha=0.4)  +
  facet_wrap(factor~., scales="free_y") +
  ggtitle("H.norm")
  

```


```{r}
W <- int.liger[["misc"]]$W
V <- int.liger[["misc"]]$V

Y.hnorm.project <- H.norm %*% (W + V$RNA)

ligerHnorm.project.umap <- uwot::umap(Y.hnorm.project, pca = 50, verbose = TRUE)

pl_hnorm.project <- as_tibble(ligerHnorm.project.umap) %>%
  mutate(cell=rownames(int.liger$misc$H.norm)) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  ggplot(aes(V1, V2, color=tech)) +
  geom_point(size=0.5, alpha=0.4)

pl_hnorm.project + ggtitle("H.norm projection")
```
```{r}
Y.W.project <- purrr::reduce(H, rbind) %*% W

ligerW.project.umap <- uwot::umap(Y.W.project, pca = 20, verbose = TRUE)

pl_w.project <- as_tibble(ligerW.project.umap) %>%
  mutate(cell=rownames(int.liger$misc$H.norm)) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  ggplot(aes(V1, V2, color=tech)) +
  geom_point(size=0.5, alpha=0.4)

pl_w.project + ggtitle("W projection")
```


How much does this projection resemble the true expression values?
```{r}
Y <- cpm(int.liger$intOut[["RNA"]])[int.features,]
Y.scaled <- scale(t(Y), scale = T, center = F) 

smp.genes <- sample(1:2000, 500)
smp.Yproject <- Y.hnorm.project[,smp.genes]
smp.Yscaled <- Y.scaled[,smp.genes]

geneX.accuracy <- melt(smp.Yproject, varnames=c('cell', "gene")) %>%
  full_join(melt(smp.Yscaled, varnames=c('cell', "gene")), by=c("cell", "gene"), suffix=c(".project", ".true")) %>%
  group_by(gene) %>%
  summarise(R2=cor(value.project, value.true, use="complete.obs"))
  # ggplot(aes(value.project, value.true)) +
  # geom_point(alpha=0.5, size=0.5) +
  # facet_wrap(gene~., scales="free") +
  # ggpubr::stat_cor()

geneX.accuracy %>% ggplot(aes(R2)) + geom_histogram(bins = 50)


```

## CCA factor model

```{r}
int.cca$misc$transfer.anchors@anchors[,"score"] %>%
  hist(breaks=100, main = "Anchor scores")

int.cca$misc$transfer.anchors@anchors[which(int.cca$misc$transfer.anchors@anchors[,"cell1"]==1),]
```


## Compare accessibility/gene expression on projections



```{r, message=FALSE}
liger.df <- 
  as_tibble(ligerHnorm.project.umap) %>%
  mutate(cell=rownames(int.liger$misc$H.norm)) %>%
  mutate(tech=int.liger$intOut@colData[cell,]) %>%
  dplyr::rename(UMAP_1=V1, UMAP_2=V2)

cca.df <- as.tibble(int.cca.seu@reductions$umap@cell.embeddings, rownames="cell") %>%
   mutate(tech=int.liger$intOut@colData[cell,])

Y.atac <- logcounts(int.cca$intOut[["ATAC"]])
Y.rna <- logcounts(int.cca$intOut[["RNA"]])

Y.impute <- Y.hnorm.project 
feature.oi <- "COL12A1"

plot.marker <- function(int.df, feature.oi){
  int.df.rna <- int.df %>% 
    left_join(as_tibble(Y.rna[feature.oi,], rownames="cell")) %>%
    left_join(as_tibble(Y.atac[feature.oi,], rownames="cell"), by="cell") %>%
    mutate(log.value=ifelse(is.na(value.x), value.y, value.x)) %>%
    group_by(tech) %>%
    dplyr::arrange(log.value) %>%
    ungroup() %>%
    filter(tech=="RNA") %>%
    ggplot(aes(UMAP_1, UMAP_2, color=log.value)) +
    geom_point(size=0.3) +
    facet_wrap(tech~.) +
    scale_color_viridis_c() +
    # scale_color_gradient2(low="", high="red") +
    ggtitle(feature.oi) +
    theme_cowplot()
  int.df.atac <- int.df %>% 
    left_join(as_tibble(Y.rna[feature.oi,], rownames="cell")) %>%
    left_join(as_tibble(Y.atac[feature.oi,], rownames="cell"), by="cell") %>%
    mutate(log.value=ifelse(is.na(value.x), value.y, value.x)) %>%
    group_by(tech) %>%
    dplyr::arrange(log.value) %>%
    ungroup() %>%
    filter(tech=="ATAC") %>%
    ggplot(aes(UMAP_1, UMAP_2, color=log.value)) +
    geom_point(size=0.3) +
    facet_wrap(tech~.) +
    scale_color_viridis_c() +
    # scale_color_gradient2(low="", high="red") +
    ggtitle(feature.oi) +
    theme_cowplot()
  ggpubr::ggarrange(int.df.rna, int.df.atac)
}


plot.marker(cca.df, "MMP2")
plot.marker(liger.df, "MMP2")
```

```{r}
plot.marker(cca.df, "GNG4")
plot.marker(liger.df, "GNG4")
```


## Explore expression-accessibility relationships
Which genes are found to have high expression-accessibility correlation?

Having a look
```{r}
normcounts(int.liger$intOut[["integrated.RNA"]]) <- t(Y.hnorm.project) ## Temporary fix

int.out <- int.liger$intOut[["integrated.RNA"]] 

imputation.query <- normcounts(int.out)[,which(int.out$tech=="ATAC")]
true.query <- cpm(int.liger$intOut[["ATAC"]])[rownames(imputation.query),] %>% as.matrix()

smp.genes <- sample(int.features, 10)
map(list(imputation.query, true.query), ~ melt(.x[smp.genes,], varnames = c("gene", "cell"))) %>%
    purrr::reduce(full_join, by=c("gene", "cell"), suffix=c(".imputed", ".true")) %>%
  ggplot(aes(value.true, log1p(value.imputed))) +
  geom_point(size=0.3) +
  facet_wrap(gene~., scales="free") +
  ggpubr::stat_cor(method = "spearman")

```

Based on model weight matrices
```{r}
## LIGER
smp.genes.liger <- melt(int.liger$misc$W, varnames=c("factor", "gene")) %>%
  top_n(10, value) %>%
  distinct(gene) %>%
  pull(gene)


map(list(imputation.query, true.query), ~ melt(.x[smp.genes.liger,], varnames = c("gene", "cell"))) %>%
    purrr::reduce(full_join, by=c("gene", "cell"), suffix=c(".imputed", ".true")) %>%
  ggplot(aes(log1p(value.true), log1p(value.imputed))) +
  geom_point(size=0.3) +
  facet_wrap(gene~., scales="free") +
  ggpubr::stat_cor(method = "pearson")
```

Plain correlation
```{r, fig.height=7, fig.width=12}

computeFeatCor <- function(int.out, int.features, reference="RNA" , query="ATAC"){
  imputation.query <- normcounts(int.out[["integrated.RNA"]])[int.features,which(int.out$tech==query)] %>% as.matrix()
  true.query <- cpm(int.out[[query]])[int.features,] %>% as.matrix()
  
  feat.cor <- map_dbl(int.features, ~ cor(imputation.query[.x,], true.query[.x,])) %>%
    set_names(int.features)
  return(feat.cor)
  }


feat.cor.liger <- computeFeatCor(int.liger$intOut, int.features)
feat.cor.cca <- computeFeatCor(int.cca$intOut, int.features)

cor_pl1 <- as.tibble(feat.cor.cca, rownames="gene") %>%
  dplyr::rename(feature.cor=value) %>%
  mutate(rank=rank(feature.cor)) %>%
  ggplot(aes(rank, feature.cor)) +
  geom_point() +
    ggrepel::geom_text_repel(data=. %>% filter(rank > (max(rank) - 20)), aes(label=gene), 
                           # xlim = c(3050, 3100),
                           # nudge_x=0.05,
                           # hjust=-1,
                           # direction="y",
                           segment.size = 0.2) +
  ggtitle("CCA") +
  ylim(-0.2, 0.6)

cor_pl2 <-
  as.tibble(feat.cor.liger, rownames="gene") %>%
  dplyr::rename(feature.cor=value) %>%
  mutate(rank=rank(feature.cor)) %>%
  ggplot(aes(rank, feature.cor)) +
  geom_point() +
  ggrepel::geom_text_repel(data=. %>% filter(rank > (max(rank) - 20)), aes(label=gene), 
                           # xlim = c(3050, 3100),
                           # nudge_x=0.05,
                           # hjust=-1,
                           # direction="y",
                           segment.size = 0.2) +
  ggtitle("LIGER") +
    ylim(-0.2, 0.6) +
    xlim(0,3100)


ggpubr::ggarrange(cor_pl1, cor_pl2)
```

```{r}
plot.marker(cca.df, "HLA-DRA")
plot.marker(liger.df, "HLA-DRA")
```
```{r}
plot.marker(cca.df, "CD74")
plot.marker(liger.df, "CD74")
```
```{r}

```










