---
title: "Joint clustering analyis"
output: html_notebook
---


```{r}
library(tidyverse)
library(Seurat)
library(SingleCellExperiment)
library(ggpubr)
library(cowplot)
source("~/multiOmic_benchmark/utils.R")

theme_set(theme_bw(base_size = 16))

## Make output directory
dir.exists("~/multiOmic_benchmark/output/")
ifelse(!dir.exists(file.path(mainDir, subDir)), dir.create(file.path(mainDir, subDir)), FALSE)
```

## Load integration output
```{r}
int.cca <- readRDS("~/models/integrateCCA_F74_SCElist_20191017.RDS")
int.liger <- readRDS("~/models/integrateliger_F74_SCElist_20191017.RDS")

int.features <- int.cca$misc$transfer.anchors@anchor.features
```

Temporary fix
```{r}
H <- int.liger[["misc"]]$H
H.norm <- int.liger[["misc"]]$H.norm

W <- int.liger[["misc"]]$W
V <- int.liger[["misc"]]$V

Y.hnorm.project <- H.norm %*% (W + V$RNA)

normcounts(int.liger$intOut[["integrated.RNA"]]) <- t(Y.hnorm.project) ## Temporary fix
```

## Co-embedding and joint clustering 

```{r}
## Make Seurat objects
int.cca.seu <- intOutput2seurat(int.cca$intOut, int.features)
int.liger.seu <- intOutput2seurat(int.liger$intOut, int.features = int.features)

## Co-embedding
int.cca.seu <- ScaleData(int.cca.seu, features = int.features)
int.cca.seu <- RunPCA(int.cca.seu)
int.cca.seu <- RunUMAP(int.cca.seu, reduction="pca", dims=1:50)

int.liger.seu <- ScaleData(int.liger.seu, features = int.features)
int.liger.seu <- RunPCA(int.liger.seu)
int.liger.seu <- RunUMAP(int.liger.seu, reduction="pca", dims=1:50)

p1 <- DimPlot(int.liger.seu, reduction = "umap", group.by="tech") + ggtitle("LIGER")
p2 <- DimPlot(int.cca.seu, reduction = "umap", group.by="tech") + ggtitle("CCA")

ggpubr::ggarrange(p1, p2, common.legend = TRUE) 
```

```{r, fig.width=14, fig.height=6}
int.seu.list <- list(liger=int.liger.seu, cca=int.cca.seu)

## Find clusters with leiden algorithm
int.seu.list <-
  map(int.seu.list, ~ FindNeighbors(.x, verbose = TRUE)) %>%
  map(~ FindClusters(.x, algorithm = 4))

# ## Extract clusters & metadata
# int.seu.list %>%
#   map(~ .x@meta.data)

## Plot
ggarrange(plotlist=imap(int.seu.list, ~ DimPlot(.x, reduction = "umap") + ggtitle(.y))) +
  ggsave("~/multiOmic_benchmark/output/")
```


## Mixing between technologies
```{r, fig.width=12, fig.height=7}
imap(int.seu.list, ~ .x@meta.data %>% mutate(method=.y)) %>%
  purrr::reduce(bind_rows) %>%
  ggplot(aes(seurat_clusters)) +
  geom_bar(aes(fill=tech)) +
  facet_grid(method~.)
```


## Agreement metric
Defined by Welch et al. 2019. Calculating how many of each cell’s nearest neighbors in the graphs built from the separate low-dimensional representations are also nearest neighbors in the graphs built from the integrated low-dimensional representations.
```{r}
## Find neighbors in single datasets 
sce.list <- as.list(int.cca$intOut[,,c("RNA", "ATAC")]@ExperimentList)

seu.list <- 
  sce.list %>% 
  map(~ as.Seurat(.x)) %>%
  map(~ ScaleData(.x, do.center=TRUE, features=int.features)) %>%
  map(~ RunPCA(.x, features=int.features))

getNNlist <- function(seurat.obj){
    seurat.obj@graphs$RNA_nn %>%
    apply(1, function(x) names(which(x==1))) %>%
    asplit(2)
  }

calculateAgreement <- function(int.seu, k=20){
  ## Find single dataset KNN graph
  seu.list <- 
    seu.list %>%
    map(~ FindNeighbors(.x, k.param = k))
  
  seu.list.nn <- map(seu.list, ~ getNNlist(.x)) 
  single.dataset.nn <- purrr::reduce(seu.list.nn, c)
  
  ## Find integrated KNN graph
  int.seu <- FindNeighbors(int.seu, k.param = k)
  nn.list.int <- getNNlist(int.seu)
  
  ## Find overlap between KNN graphs
  common_nns <- map(names(single.dataset.nn), ~ sum(nn.list.int[[.x]] %in% single.dataset.nn[[.x]])) %>% purrr::reduce(sum) 
  agreement.score <- {common_nns/(k*length(single.dataset.nn))}
  return(agreement.score)
  }


liger.agr <- map(seq(5,40, by=5), ~ calculateAgreement(int.seu.list$liger, k=10))
cca.agr <- map(seq(5,40, by=5), ~ calculateAgreement(int.seu.list$cca, k=10))

```

```{r}

```

## Purity of annotated cell types in clusters 





