---
title: "Joint clustering analyis"
output: html_notebook
---


```{r}
library(tidyverse)
library(Seurat)
library(SingleCellExperiment)
library(ggpubr)
library(cowplot)
library(reshape2)
source("~/multiOmic_benchmark/utils.R")

theme_set(theme_bw(base_size = 16))

## Make output directory
outdir <- "~/multiOmic_benchmark/output/20191024_JointClustering/"
ifelse(!dir.exists(outdir), dir.create(outdir), FALSE)
```

## Load integration output
```{r}
int.cca <- readRDS("~/models/integrateCCA_F74_SCElist_20191017.RDS")
int.liger <- readRDS("~/models/integrateliger_F74_SCElist_20191017.RDS")

int.features <- int.cca$misc$transfer.anchors@anchor.features
```

Temporary fix for liger output 
```{r}
H <- int.liger[["misc"]]$H
H.norm <- int.liger[["misc"]]$H.norm

W <- int.liger[["misc"]]$W
V <- int.liger[["misc"]]$V

Y.hnorm.project <- H.norm %*% (W + V$RNA)

normcounts(int.liger$intOut[["integrated.RNA"]]) <- t(Y.hnorm.project) ## Temporary fix
```

## Co-embedding and joint clustering 

```{r}
## Make Seurat objects
int.cca.seu <- intOutput2seurat(int.cca$intOut, int.features)
int.liger.seu <- intOutput2seurat(int.liger$intOut, int.features = int.features)

## Co-embedding
int.cca.seu <- ScaleData(int.cca.seu, features = int.features)
int.cca.seu <- RunPCA(int.cca.seu)
int.cca.seu <- RunUMAP(int.cca.seu, reduction="pca", dims=1:50)

int.liger.seu <- ScaleData(int.liger.seu, features = int.features)
int.liger.seu <- RunPCA(int.liger.seu)
int.liger.seu <- RunUMAP(int.liger.seu, reduction="pca", dims=1:50)

p1 <- DimPlot(int.liger.seu, reduction = "umap", group.by="tech") + ggtitle("LIGER")
p2 <- DimPlot(int.cca.seu, reduction = "umap", group.by="tech") + ggtitle("CCA")

ggpubr::ggarrange(p1, p2, common.legend = TRUE) 
```

```{r, fig.width=14, fig.height=6}
int.seu.list <- list(liger=int.liger.seu, cca=int.cca.seu)

## Find clusters with leiden algorithm
int.seu.list <-
  map(int.seu.list, ~ FindNeighbors(.x, verbose = TRUE)) %>%
  map(~ FindClusters(.x, algorithm = 4, resolution = 1))

## Plot
ggarrange(plotlist=imap(int.seu.list, ~ DimPlot(.x, reduction = "umap") + ggtitle(.y))) +
  ggsave(paste0(outdir, "UMAP_clustering.png"), height=6, width=12)
```


## Mixing between technologies
```{r, fig.width=12, fig.height=7, message=FALSE, warning=FALSE}
imap(int.seu.list, ~ .x@meta.data %>% mutate(method=.y)) %>%
  purrr::reduce(bind_rows) %>%
  ggplot(aes(seurat_clusters)) +
  geom_bar(aes(fill=tech)) +
  facet_grid(method~.) +
  ggsave(paste0(outdir, "clusters_by_tech.png"), height=6, width=12)
```


## Agreement metric
Defined by Welch et al. 2019. Calculating how many of each cell’s nearest neighbors in the graphs built from the separate low-dimensional representations are also nearest neighbors in the graphs built from the integrated low-dimensional representations.
```{r, message=FALSE, warning=FALSE, fig.height=10, fig.width=14}
## Find neighbors in single datasets 
sce.list <- as.list(int.cca$intOut[,,c("RNA", "ATAC")]@ExperimentList)

seu.list <- 
  sce.list %>% 
  map(~ as.Seurat(.x)) %>%
  map(~ ScaleData(.x, do.center=TRUE, features=int.features)) %>%
  map(~ RunPCA(.x, features=int.features))

getNNlist <- function(seurat.obj){
    seurat.obj@graphs$RNA_nn %>%
    apply(1, function(x) names(which(x==1))) %>%
    asplit(2)
  }

calculateAgreement <- function(int.seu, tech, k=20){
  ## Find single dataset KNN graph
  single.seu <- FindNeighbors(seu.list[[tech]], k.param = k)
  single.nn <- getNNlist(single.seu)  
  # single.dataset.nn <- purrr::reduce(seu.list.nn, c)
  
  ## Find integrated KNN graph (each dataset separately)
  int.seu <- FindNeighbors(int.seu[,int.seu$tech==tech], k.param = k)
  nn.list.int <- getNNlist(int.seu)

  ## Find overlap between KNN graphs
  # common_nns <- map(names(single.dataset.nn), ~ sum(nn.list.int[[.x]] %in% single.dataset.nn[[.x]])) %>% purrr::reduce(sum) 
  common_nns <- map_dbl(names(single.nn), ~ sum(nn.list.int[[.x]] %in% single.nn[[.x]])) %>%
      set_names(names(single.nn))
  agreement.score <- {sum(common_nns)/(k*(length(single.nn)))}
  return(list(common_nns, agreement.score))
  }


liger.agr.atac <- map(seq(5,40, by=5), ~ calculateAgreement(int.seu.list$liger, tech="ATAC", k=.x))
cca.agr.atac <- map(seq(5,40, by=5), ~ calculateAgreement(int.seu.list$cca, tech="ATAC", k=.x))

liger.agr.rna <- map(seq(5,40, by=5), ~ calculateAgreement(int.seu.list$liger, tech="RNA", k=.x))
cca.agr.rna <- map(seq(5,40, by=5), ~ calculateAgreement(int.seu.list$cca, tech="RNA", k=.x))

# liger.agr.k20 <- calculateAgreement(int.seu = int.seu.list$liger, k=50)
# cca.agr.k20 <- calculateAgreement(int.seu = int.seu.list$cca, k=50)
# 
# pl1 <- data.frame(sharedNNs_per_cell = liger.agr.k20[[1]], cell=names(liger.agr.k20[[1]])) %>%
#   mutate(tech=ifelse(str_detect(cell, "_"), "RNA", "ATAC")) %>%
#   ggplot(aes(sharedNNs_per_cell, fill=tech)) + geom_histogram(bins=50) +
#   facet_wrap(tech~.) +
#     xlim(0,50) +
#   ggtitle("LIGER (k=20)",subtitle = paste("Agreement score =", round(liger.agr.k20[[2]], 2))) 
# 
# pl2 <- data.frame(sharedNNs_per_cell = cca.agr.k20[[1]], cell=names(cca.agr.k20[[1]])) %>%
#   mutate(tech=ifelse(str_detect(cell, "_"), "RNA", "ATAC")) %>%
#   ggplot(aes(sharedNNs_per_cell, fill=tech)) + geom_histogram(bins=50) +
#   facet_wrap(tech~.) +
#   xlim(0,50) +
#   ggtitle("CCA (k=20)",subtitle = paste("Agreement score =", round(cca.agr.k20[[2]], 2)))
# 
# ggarrange(pl1, pl2, nrow=2) +
#   ggsave(paste0(outdir, "agreement_k20.png"), height = 8, width=8)
# pl1

  
pl_atac <- bind_rows(
  imap_dbl(liger.agr.atac, ~ .x[[2]]) %>% set_names(seq(5,40, by=5)) %>% data.frame(agr=., k=as.double(names(.))) %>% mutate(method="LIGER"),
  imap_dbl(cca.agr.atac, ~ .x[[2]]) %>% set_names(seq(5,40, by=5)) %>% data.frame(agr=., k=as.double(names(.))) %>% mutate(method="CCA")
  ) %>%
  ggplot(aes(as.numeric(k), agr, color=method)) +
  geom_point(size=2) +
  ggtitle("ATAC")

pl_rna <- bind_rows(
  imap_dbl(liger.agr.rna, ~ .x[[2]]) %>% set_names(seq(5,40, by=5)) %>% data.frame(agr=., k=as.double(names(.))) %>% mutate(method="LIGER"),
  imap_dbl(cca.agr.rna, ~ .x[[2]]) %>% set_names(seq(5,40, by=5)) %>% data.frame(agr=., k=as.double(names(.))) %>% mutate(method="CCA")
  ) %>%
  ggplot(aes(as.numeric(k), agr, color=method)) +
  geom_point(size=2) +
  ggtitle("RNA")

ggarrange(pl_rna, pl_atac, ncol=1, nrow=2)
```



## Purity of annotated cell types in clusters 
```{r, fig.width=8, fig.height=6}
annotation.df <- read.csv("~/my_data/F74_RNA_obs.csv")
annotation.df <- annotation.df %>%
  mutate(cell=str_remove(as.character(X), "F74_1_") %>% str_c(ifelse(batch==0,'_1', "_2"))) 

cell.type.pal <- brewer_palette_4_values(unique(annotation.df$annotation), palette = "Accent")

int.liger.seu@meta.data %<>%
  rownames_to_column("cell") %>%
  left_join(annotation.df, by="cell") 
int.liger.seu@meta.data %<>% column_to_rownames("cell")

int.cca.seu@meta.data %<>%
  rownames_to_column("cell") %>%
  left_join(annotation.df, by="cell") 
int.cca.seu@meta.data %<>% column_to_rownames("cell")

DimPlot(int.liger.seu, reduction = "umap", group.by="annotation") + ggtitle("LIGER")
DimPlot(int.cca.seu, reduction = "umap", group.by="annotation")  + ggtitle("CCA")
```

### Fraction of cells per cell type in each cluster
Fraction is calculated on RNA cells only
```{r, fig.width=12, fig.height=10, warning=FALSE, message=FALSE}
imap(int.seu.list, ~ .x@meta.data %>%
       rownames_to_column("cell") %>%
       left_join(annotation.df, by="cell") %>%
        filter(tech=="RNA") %>%
       column_to_rownames("cell") %>%
       group_by(seurat_clusters) %>%
       mutate(n_clusters=n()) %>%
       ungroup() %>%
       group_by(annotation, seurat_clusters) %>%
       mutate(count=n()) %>%
       ungroup() %>%
       mutate(frac=count/n_clusters) %>%
       filter(!is.na(annotation)) %>%
        ggplot(aes(y=annotation, x=as.factor(seurat_clusters), color=frac, size=frac)) +
       geom_point() +
       scale_color_gradient(low="grey70", high="red", name="Fraction x cluster") +
       #  geom_bar(position="fill") +
        ggtitle(.y) 
       # coord_flip() +
       # scale_color_brewer()
  )



```

### Fraction of cells per cluster of each cell type
Visualize cell type purity of clusters from integrated datasets (fraction of cells per cell type in each cluster)
```{r, fig.width=7, fig.height=8, warning=FALSE, message=FALSE}
make_annoXcluster_frac <- function(int.seu){
  anno.frac.mat <- int.seu@meta.data %>%
         rownames_to_column("cell") %>%
         left_join(annotation.df, by="cell") %>%
         column_to_rownames("cell") %>%
         filter(tech=="RNA") %>%
         group_by(annotation) %>%
         mutate(n_anno=n()) %>%
         ungroup() %>%
         group_by(annotation, seurat_clusters) %>%
         mutate(count=n()) %>%
         ungroup() %>%
         mutate(frac=count/n_anno) %>%
         select(seurat_clusters, annotation, frac) %>%
         distinct() %>%
         spread(seurat_clusters, frac) %>%
         column_to_rownames('annotation') %>%
         as.matrix()
  return(anno.frac.mat)  
}

make_annoXcluster_frac(int.seu.list$liger) %>%     
  {ifelse(is.na(.), 0, .)} %>%
  pheatmap::pheatmap(color = c('white',brewer.pal(15,"Reds")), main = "LIGER")
make_annoXcluster_frac(int.seu.list$cca) %>%     
  {ifelse(is.na(.), 0, .)} %>%
  pheatmap::pheatmap(color = c('white',brewer.pal(15,"Reds")), main = "CCA")


```



Calculate cluster purity:

- Find most abundant cell type per cluster
- calculate fraction of cells of that cell type in the cluster (excluding ATAC cells w no assignment)

```{r, fig.width=15, fig.height=5}
clusterPurity <- function(meta.data){
  meta.data %>%
       rownames_to_column("cell") %>%
       left_join(annotation.df, by="cell") %>%
       column_to_rownames("cell")  %>%
    # filter(!is.na(annotation)) %>%
    filter(tech=="RNA") %>%
    select(tech, seurat_clusters, annotation) %>%
    group_by(seurat_clusters, annotation) %>%
    mutate(n_cell_type=n()) %>%
    ungroup() %>%
    group_by(seurat_clusters) %>%
    summarise(n_cluster=n(),
              cluster.cell.type=annotation[max(n_cell_type)],
              n_cell_type=max(n_cell_type)) %>%
    mutate(frac=n_cell_type/n_cluster)
}

pl1 <- clusterPurity(int.seu.list$cca@meta.data) %>%
  arrange(-frac) %>%
  mutate(seurat_clusters=factor(seurat_clusters, levels=unique(seurat_clusters))) %>%
  ggplot(aes(seurat_clusters, frac, fill=cluster.cell.type)) +
  geom_col() +
  scale_fill_manual(values = cell.type.pal) +
  ggtitle("CCA")

pl2 <- clusterPurity(int.seu.list$liger@meta.data) %>%
  arrange(-frac) %>%
  mutate(seurat_clusters=factor(seurat_clusters, levels=unique(seurat_clusters))) %>%
  ggplot(aes(seurat_clusters, frac, fill=cluster.cell.type)) +
  geom_col() +
  scale_fill_manual(values = cell.type.pal) +
  ggtitle("LIGER")


ggarrange(pl1, pl2, common.legend = TRUE)
```

### Find markers in joint clusters impure clusters
```{r, fig.height=15, fig.width=15}
clusterPurity(int.seu.list$cca@meta.data) %>% filter(frac < 0.5) %>% pull(seurat_clusters) %>% as.numeric()
# FindMarkers(int.seu.list$cca, ident.1 = 8)
# 
# int.seu.list$liger <- FindAllMarkers(int.seu.list$liger, min.pct = 0.2, logfc.threshold = 0.3, only.pos = TRUE)

cl4_markers <- FindMarkers(int.seu.list$liger, ident.1 = '4')
marks <- cl4_markers %>%
  rownames_to_column("gene") %>%
  arrange(p_val_adj) %>%
  .[1:10,] %>%
  pull(gene)

FeaturePlot(int.seu.list$liger, features = marks) 
```


<!-- Calculate purity w different clustering modularity -->
<!-- ```{r} -->
<!-- for (res in seq(0.2, 1, 0.2)) { -->
<!--   print(paste("Resolution=", res)) -->
<!--   int.seu.list <- map(int.seu.list, ~ FindClusters(.x, algorithm = 4, resolution = res)) -->
<!-- } -->

<!-- int.seu.list$liger@meta.data %>% -->
<!--   rownames_to_column("cell") %>% -->
<!--   pivot_longer(cols = str_subset(colnames(int.seu.list$liger@meta.data),"snn"), names_to = "res", values_to = "cluster") %>% -->
<!--   mutate(res=as.numeric(str_remove(res,"RNA_snn_res."))) %>% -->
<!--   select(-seurat_clusters) %>% -->
<!--   dplyr::rename(seurat_clusters = cluster) %>% -->
<!--   split(.$res) %>% -->
<!--   imap( ~ column_to_rownames(.x, "cell")) %>% -->
<!--   imap( ~ clusterPurity(.x) %>% mutate(res=.y)) %>% -->
<!--   purrr::reduce(bind_rows) %>% -->
<!--   ggplot(aes(res, frac)) + -->
<!--   geom_violin() + -->
<!--   geom_boxplot(width=0.2) + -->
<!--   geom_jitter()  -->

<!-- int.seu.list$cca@meta.data %>% -->
<!--   rownames_to_column("cell") %>% -->
<!--   pivot_longer(cols = str_subset(colnames(int.seu.list$liger@meta.data),"snn"), names_to = "res", values_to = "cluster") %>% -->
<!--   mutate(res=as.numeric(str_remove(res,"RNA_snn_res."))) %>% -->
<!--   select(-seurat_clusters) %>% -->
<!--   dplyr::rename(seurat_clusters = cluster) %>% -->
<!--   split(.$res) %>% -->
<!--   imap( ~ column_to_rownames(.x, "cell")) %>% -->
<!--   imap( ~ clusterPurity(.x) %>% mutate(res=.y)) %>% -->
<!--   purrr::reduce(bind_rows) %>% -->
<!--   ggplot(aes(res, frac)) + -->
<!--   geom_violin() + -->
<!--   geom_boxplot(width=0.2) + -->
<!--   geom_jitter()  -->



<!-- ``` -->


