---
title: "SnapATAC output EDA"
output: html_notebook
---


```{r}
library(SnapATAC)
library(Seurat)
```


Save files
```{r}
saveRDS(x.sp, file = "~/my_data/cellranger-atac110_count_30439_WSSS8038360_GRCh38-1_1_0.snapATAC.RDS")
x.sp <- readRDS( "~/my_data/cellranger-atac110_count_30439_WSSS8038360_GRCh38-1_1_0.snapATAC.RDS")
```


Repeat visualization on gmat
```{r}
## Make suerat object
f74.seu <- snapToSeurat(
    obj=x.sp, 
    eigs.dims=1:20, 
    norm=TRUE,
    scale=TRUE
    )

## Save UMAP based on SnapATAC processing
f74.seu <- RunUMAP(f74.seu, reduction = "SnapATAC", dims=signif.dims, reduction.name = "umap.snap")


f74.seu <- FindVariableFeatures(f74.seu)
f74.seu <- RunPCA(f74.seu, dims=1:50)
f74.seu <- RunUMAP(f74.seu, reduction = "pca", dims=1:50, verbose=FALSE) 

Seurat::Idents(f74.seu) <- x.sp@cluster

ggpubr::ggarrange(
  DimPlot(f74.seu, reduction = "umap") + ggtitle("gmat"),
  DimPlot(f74.seu, reduction = "umap.snap") + ggtitle("bmat - snap") 
  )

```


```{r}
plotly::ggplotly(DimPlot(f74.seu, reduction = "umap", group.by = NULL) + ggtitle("gmat"))
```

```{r, fig.height=15, fig.width=15}
FeaturePlot(f74.seu, features = c('PTPRC','CD4','CD8A','CD8B','CD79A','FOXN1','EPCAM','PDGFRA','GNG4', 'FOXP3','RAG1','RAG2','NKG7','CCR7'), reduction = "umap.snap")
```

```{r, fig.height=15, fig.width=15}
DimHeatmap(f74.seu, dims = 1:6, balanced = TRUE,ncol = 1)
```

### Differential analysis 
<!-- Snap ATAC way -->
<!-- ```{r} -->
<!-- cluster = 1 -->
<!-- DARs= findDAR( -->
<!--     obj=x.sp, -->
<!--     input.mat="gmat", -->
<!--     cluster.pos=cluster, -->
<!--     # cluster.neg = 5, -->
<!--     cluster.neg.method="random", -->
<!--     test.method="exactTest", -->
<!--     bcv=0.4, #0.4 for human, 0.1 for mouse -->
<!--     seed.use=42 -->
<!--   ); -->


<!-- gmat.cl <- x.sp@gmat[which(x.sp@cluster==cluster),] -->
<!-- gmat.neg <- x.sp@gmat[which(x.sp@cluster!=cluster),] -->

<!-- hist(as.matrix(gmat.cl), breaks=50) -->


<!-- # DARs %>% -->
<!--   # mutate(gene=colnames(x.sp@gmat)) %>% -->
<!--   # arrange() %>% -->
<!--   # filter(PValue < 0.05) -->

<!-- hist(DARs$PValue, breaks = 20) -->

<!-- DARs$FDR = p.adjust(DARs$PValue, method="BH") -->
<!-- idy = which(DARs$FDR < 5e-2 & DARs$logFC > 0) -->
<!-- plot(DARs$logCPM, DARs$logFC,  -->
<!--     pch=19, cex=0.1, col="grey",  -->
<!--     ylab="logFC", xlab="logCPM", -->
<!--     main=paste("Cluster", cluster) -->
<!--   ) -->
<!-- points(DARs$logCPM[idy],  -->
<!--     DARs$logFC[idy],  -->
<!--     pch=19,  -->
<!--     cex=0.5,  -->
<!--     col="red" -->
<!--   ) -->

<!-- ``` -->
```{r}
DAGs_cl <- FindAllMarkers(f74.seu, min.pct = 0.2, logfc.threshold = 0.3)
DAGs_cl <- 
  DAGs_cl %>%
  mutate(positive=avg_logFC > 0)
```
```{r, fig.height=14, fig.width=9}
top10_markers <- DAGs_cl %>%
  filter(positive) %>%
  group_by(cluster) %>%
  arrange(p_val_adj) %>%
  top_n(10) %>%
  pull(gene) %>%
  unique()

top10_markers <- DAGs_cl %>%
  filter(positive) %>%
  group_by(cluster) %>%
  # arrange(p_val_adj) %>%
  top_n(10,wt = - p_val_adj) %>%
  pull(gene)

DotPlot(f74.seu, features = top10_markers) +
  coord_flip()
```

Expression of markers in RNA data

```{r, fig.width=12, fig.height=16}
plot_marker_ATACvsRNA <- function(marker){
    ggpubr::ggarrange(
      FeaturePlot(f74.seu, feature=marker, reduction="umap.snap") + ggtitle("ATAC"), FeaturePlot(f74.rna.seu, feature=marker) + ggtitle("RNA")
    ) %>% annotate_figure(top=marker)
  }

plot_cluster_marker_ATACvsRNA <- function(cl){
  top10_cluster <-
    DAGs_cl %>%
    filter(positive) %>%
    filter(cluster==cl) %>%
    filter(gene %in% rownames(f74.rna.seu)) %>%
    top_n(10,wt = - p_val_adj) %>%
    pull(gene)
  ggarrange(plotlist = map(top10_cluster, ~ plot_marker_ATACvsRNA(.x)), ncol=2, nrow=5) %>% 
    annotate_figure(fig.lab=paste("Cluster", cl), fig.lab.size = 20, fig.lab.face = "bold")
  }

map(1:10, ~ plot_cluster_marker_ATACvsRNA(.x))
```

## Differential analysis only between the putative T-cell clusters
```{r, fig.width=8, fig.height=15}
t_clusters <- c("1", '2', '3', '4', '5')
DAGs_tcells <- map(t_clusters, ~ FindMarkers(f74.seu, ident.1 = .x, ident.2 = setdiff(t_clusters,.x), min.pct = 0.2, logfc.threshold = 0.3))

top20_tcells_markers <- imap(DAGs_tcells, ~ .x %>% rownames_to_column() %>% filter( avg_logFC > 0) %>% mutate(cluster=.y)) %>%
  purrr::reduce(bind_rows) %>%
  # filter(positive) %>%
  group_by(cluster) %>%
  top_n(20,wt = - p_val_adj) %>%
  pull(rowname)


DotPlot(f74.seu, features = top20_tcells_markers) +
  coord_flip() +
  ggtitle("T-cell subtypes markers (DA between clusters 1-5)")
```

Only between cluster 1 and 2
```{r}
DAGs_tcells_12 <- FindMarkers(f74.seu, ident.1 = '1', ident.2 = '2', min.pct = 0.2, logfc.threshold = 0.3)

top20_cl12_markers <- DAGs_tcells_12 %>% 
  rownames_to_column() %>% 
  filter( avg_logFC > 0) %>% 
  top_n(20,wt = - p_val_adj) %>%
  pull(rowname)

DotPlot(f74.seu, features = top20_cl12_markers) +
  coord_flip() +
  ggtitle("DA between clusters 1-2")
```



```{r}
plotly::ggplotly(DimPlot(f74.rna.seu, group.by = "annotation"))
```


## Embedding on most variable genes in ATAC (not in gene expression)
```{r}
DefaultAssay(f74.seu) <- "ACTIVITY"
f74.seu.hvgatac <- FindVariableFeatures(f74.seu)
f74.seu.hvgatac <- RunPCA(f74.seu.hvgatac, dims=1:30)
f74.seu.hvgatac <- RunUMAP(f74.seu.hvgatac, reduction = "pca", dims=1:30, verbose=FALSE) 

f74.seu.hvgatac <- AddMetaData(f74.seu.hvgatac, x.sp@cluster, col.name='SnapATAC_cluster')
DimPlot(f74.seu.hvgatac, reduction = "umap", group.by = "SnapATAC_cluster") 
```

## Save markers
```{r}
DAGs_cl %>% filter(positive) %>% write_csv("~/my_data/integrated_thymus/F74_ATAC_clusterMarkers_allVSall_20191104.csv")
imap(DAGs_tcells, ~ .x %>% 
  rownames_to_column("gene") %>%
  filter(avg_logFC > 0) %>%
  mutate(cluster=.y)) %>%
  purrr::reduce(bind_rows) %>%
  write_csv("~/my_data/integrated_thymus/F74_ATAC_clusterMarkers_1to5VS1to5_20191104.csv")

DAGs_tcells_12 %>% 
  rownames_to_column("gene") %>%
  write_csv("~/my_data/integrated_thymus/F74_ATAC_clusterMarkers_1VS2_20191104.csv")
```



## Thoughts

- There seems to be a correspondance for gene expression and accessibility in many "accessibility markers" (see Cluster 6) --> maybe integration would work better if I don't take only the most variable features in the RNA data. 


